Article Title

### 2013-12-15 plos.tex
minor

\section*{Introduction}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been shown to have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011,Ecker:2011}. In addition, neural correlations are hypothesized to reflect key features of functional connectivity in neural circuits.  Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus tuning similarity \cite{Kohn:2005,Ko:2011}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.

@@

Yet, a complete correlation matrix provides more information than the equivalent number of isolated pairwise correlations. In early studies, the effects of correlations were extrapolated from isolated pairs to entire populations through simulation and theoretical analysis \cite{Shadlen:1998,Zohary:1994}. In contrast, modern multineuronal recordings from increasingly large and dense populations of highly interconnected neurons allow estimation of various low-dimensional representations of the entire correlation matrix, which can enable inquiry into the nature of underlying physiological mechanisms and computational principles.

@@

Alternatively, the correlation matrix can be formulated through the partial correlations between pairs of neurons. The partial correlation between a pair of neurons is the linear correlation remaining after accounting for correlations to all the other neurons. Partial correlations carry particular significance when conditional dependencies between all the variables are thought to be dominated by linear effects. In such systems, zero  partial correlation between two variables implies their conditional independence.  Conditional independence between a pair of variables suggests a lack of direct interaction between the underlying processes. Therefore, estimation of networks of partial correlations (sometimes called \emph{association networks}) have been used to infer interactions between components in complex systems such as gene interaction networks \cite{Schafer:2005,Peng:2009} and functional connectivity between brain regions in fMRI signals \cite{Varoquaux:2012,Ryali:2012}. In \emph{elliptical distributions} (of the form $g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial interactions express conditional dependencies. A multivariate normal distribution specified through its graph of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}. With increasing departure from linearity, the correspondence between conditional dependence and partial correlations diminishes or completely breaks down \cite{Loh:2012}. For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations.

In this study, we pursued two related aims: (a)~improved estimation of neural correlation matrices and (b)~discovery of low-dimensional structure of correlations in recordings of multineuronal activity to facilitate their interpretation.

@@

@@

Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward a low-dimensional approximation referred to here as the \emph{target estimate} \cite{Schafer:2005,Bickel:2006}.  The usual estimate $C_{\sf 0}$ has the advantage of being unbiased, \emph{i.e.}\;$\E{C_0}=\Sigma$ but, on average, it falls far from $\Sigma$ due to its sensitivity to sampling noise in the data.  Low-dimensional estimates of various forms are typically less susceptible to sampling noise but are also liable to introduce their respective biases away from truth $\Sigma$.  Regularization works by striking a favorable balance between bias and variability. Regularization can produce \emph{some} improvement even with an arbitrary target estimate that has no relation to the true dependencies in the data.  Yet, when the target estimate is well suited for capturing the important features of the true covariance matrix with few terms, it will introduce minimal bias and outperform other estimators.

Figure \ref{fig:01} illustrates the challenges of estimating the correlation matrix from a finite sample. Panel \ref{fig:01}\,A depicts the sample correlation matrix $R_{\sf 0}$ of a period of somatic calcium signals of a group of 298 neurons in mouse visual cortex. $R_{\sf 0}$ is computed by normalizing the covariance matrix $C_{\sf 0}$ by the sample variances:
\begin{equation}
    R_{\sf 0} = \left(I\circ C_{\sf 0}\right)^{-\frac 1 2} C_{\sf 0} \left(I\circ C_{\sf 0}\right)^{-\frac 1 2}
\end{equation}
Here $\circ$ denotes entrywise matrix product (Hadamard product). Thus $I\circ C_{\sf 0}$ represents the diagonal matrix with the diagonal elements from $C_{\sf 0}$, \emph{i.e.} the sample variances. 
Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are much higher than typical firing rate correlations on shorter temporal scales. To show greater detail, panel \ref{fig:01}\,B magnifies the fragment of the sample correlation matrix that is outlined in \ref{fig:01}\,A. The matrix of partial correlations $P$ is obtained by scaling the inverse of the correlation matrix:
\begin{equation}
    P = -D^{-1/2} R^{-1} D^{-1/2}
\end{equation}
where $D=I\circ R^{-1}$ is the diagonal matrix with the diagonal elements of the inverse of the correlation matrix $R$. Thus $P$ contains $-1$s on its diagonal and the off-diagonal elements contain the pairwise partial correlations. Figure \ref{fig:01}\,C shows the magnified fragment of the partial correlation matrix $P$.

Panels \ref{fig:01}\,D and E depict a different estimate obtained from the same data by \emph{covariance selection} \cite{Dempster:1972}. In covariance selection, only a subset of partial correlations is fitted while the rest are fixed at zero. The resulting estimate closely reproduces the sample correlation matrix (Fig.~\ref{fig:01}\,D), but its partial correlation structure appears substantially different(Fig.~\ref{fig:01}\,E). In this particular example  31,501 of the 44,253 (71.2\%) total partial correlations are set to zero. This alternative estimate is attractive for a few reasons. First, it has fewer free parameters and is therefore less susceptible to sampling noise. Additionally, assuming that partial correlations do indeed reflect elements of the functional connectivity, the second estimate also appears more physiologically plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.

@@

Estimation of the correlation matrix relies on estimating the means and variances of the activity of individual cells.  To estimate the mean activity, we resort to the usual \emph{sample mean} without attempting to improve it. Estimation of a \emph{covariance matrix} is equivalent to estimating both  variances and correlations. Therefore, we rephrase the problem as one of optimal estimation of the covariance matrix given an estimate of the mean firing rates.

@@

\section*{Discussion}
\subsection*{Big neural data}
\subsection*{Meaning of interactions}
\subsection*{Fitting probabilistic models (\emph{e.g.}\;GLMs) vs.\;covariance estimation.}
\subsection*{Future directions}

>>>>

\section*{Introduction}
\subsection*{Neural correlations}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been shown to have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011,Ecker:2011}. In addition, neural correlations are hypothesized to reflect key features of functional connectivity in neural circuits.  Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus tuning similarity \cite{Kohn:2005,Ko:2011}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.

@@

\subsection*{Additional information in neural correlation matrices}
Yet, a complete correlation matrix provides more information than the equivalent number of isolated pairwise correlations. In early studies, the effects of correlations were extrapolated from isolated pairs to entire populations through simulation and theoretical analysis \cite{Shadlen:1998,Zohary:1994}. In contrast, modern multineuronal recordings from increasingly large and dense populations of highly interconnected neurons allow estimation of various low-dimensional representations of the entire correlation matrix, which can enable inquiry into the nature of underlying physiological mechanisms and computational principles.

@@

Alternatively, the correlation matrix can be formulated through the partial correlations between pairs of neurons. The partial correlation between a pair of neurons is the linear correlation remaining after accounting for correlations to all the other neurons. Partial correlations carry particular significance when conditional dependencies between all the variables are thought to be dominated by linear effects. In such systems, zero  partial correlation between two variables implies their conditional independence.  Conditional independence between a pair of variables suggests a lack of direct interaction between the underlying processes. Therefore, estimation of networks of partial correlations (sometimes called \emph{association networks}) have been used to infer interactions between components in complex systems such as gene interaction networks \cite{Schafer:2005,Peng:2009} and functional connectivity between brain regions in fMRI signals \cite{Varoquaux:2012,Ryali:2012}. In \emph{elliptical distributions} (those of the form $\frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is a normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial interactions express conditional dependencies. A multivariate normal distribution specified through its graph of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}. With increasing departure from linearity, the correspondence between conditional dependence and partial correlations diminishes or completely breaks down \cite{Loh:2012}. For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations.

\subsection*{Estimation of neural correlation matrices}
In this study, we pursued two related aims: (a)~improved estimation of neural correlation matrices and (b)~discovery of low-dimensional structure of correlations in recordings of multineuronal activity to facilitate their interpretation.

@@

Given a covariance matrix estimate $C$, the correlation matrix $R$ and partial correlation matrix $P$ are computed as
\begin{equation}
    R = \left(I\circ C\right)^{-\frac 1 2} C \left(I\circ C\right)^{-\frac 1 2}
\end{equation}
\begin{equation}
    P = -\left(I\circ C^{-1}\right)^{-\frac 1 2} C^{-1} \left(I\circ C^{-1}\right)^{-\frac 1 2}
\end{equation}
Here $\circ$ denotes entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix. $R$ contains 1s on its diagonal whereas $P$ contains $-1$s on the diagonal. The off-diagonal elements of $P$ contain the partial correlations between corresponding pairs of neurons.

@@

Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward a low-dimensional approximation \cite{Schafer:2005,Bickel:2006}.  The usual estimate $C_{\sf 0}$ has the advantage of being unbiased ($\E{C_0}=\Sigma$) but, on average, it falls far from $\Sigma$ due to its sensitivity to sampling noise in the data.  Low-dimensional estimates of various forms are typically less susceptible to sampling noise but are also liable to introduce their respective biases away from truth $\Sigma$.  Regularization works by striking a favorable balance between bias and variability. Regularization can produce \emph{some} improvement even with an arbitrary target estimate that has no relation to the true dependencies in the data.  Yet, when the target estimate is well suited for capturing the important features of the true covariance matrix with few terms, it will introduce minimal bias and outperform other estimators.

Figure \ref{fig:01} illustrates the challenges of estimating the correlation matrix from a finite sample and one example of regularization.  Panel A depicts the correlation matrix computed from the sample covariance matrix $C_{\sf 0}$ for a period of somatic calcium signals of a group of 298 neurons in mouse visual cortex. To show detail, panel B magnifies the fragment outlined in panel A. Panel C shows the same fragment from the partial correlation matrix also computed from $C_{\sf 0}$.  Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales.

Panels \ref{fig:01}\,D and E depict the correlation matrix and the partial correlation matrix computed from a different estimate, which we will refer to here as $C_{\sf sparse}$, which is computed through \emph{covariance selection} \cite{Dempster:1972}. In covariance selection, only a subset of partial correlations is fitted while the rest are fixed at zero. The resulting estimate closely reproduces the sample correlation matrix (Fig.~\ref{fig:01}\,D), but its partial correlation structure appears substantially different(Fig.~\ref{fig:01}\,E). In this particular example  31,501 of the 44,253 (71.2\%) total partial correlations are set to zero. This alternative estimate is attractive for a few reasons. First, it has fewer free parameters and is therefore less susceptible to sampling noise. Additionally, assuming that partial correlations do indeed reflect elements of the functional connectivity, the second estimate also appears more physiologically plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.

@@

@@

\section*{Discussion}
\subsection*{Functional connectivity from big neural data}
Breakthroughs in neuroscience

\subsection*{Meaning of interactions}
\subsection*{Fitting probabilistic models (\emph{e.g.}\;GLMs) vs.\;covariance estimation.}
\subsection*{Future directions}
